\textbf{Modelling Neuroplasticity: Using Machine Learning to Learn About TMS Data}

### 2023-04-14 paper/paper.tex
Paper

@@

The distribution of the evoked potentials was always exponential. However, the
$\beta$ parameter of the distribution varies across subject groups and session
types, as shown in Figures \ref{fig:figure1} and \ref{fig:figure2}.

@@

    \begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|  }
    \hline
    &       Mean & Median \\
    \hline
    HC BL & 196.77 & 149.36 \\
    \hline
    HC SWD & 165.98 & 99.76 \\
    \hline \\ 
    MDD BL & 187.56 & 127.69\\ 
    \hline 
    MDD SWD & 247.57 & 151.07 \\
    \hline
    \end{tabular}

@@

Each of the means in the above table corresponds to the estimated $\beta$
parameter of the distribution of the paired pulses on each subject group.

@@

In order to test these measures, we look at whether they improve the performance
of a ML classification model and in what degree.\footnote{We use a
classification model for practical reasons ---namely, because the samples in our
data were conveniently labeled by subject group and sleep session.} Our case
study is a random forest model in Python tasked with classifying every
observation with the appropriate subject label. In other words, its goal is to
infer, based on the properties of each TMS response, the group of the subject
upon which the stimulation was elicited, as well as the type of sleep session
after which it occurred. Our hypothesis is that pulse-specific measures of relative
amplitude will significantly improve a random forest's performance.

@@

When trained on the raw data ---that is, without the inclusion of pulse-specific
MRA--- the model's accuracy was of $\approx34.2 \%$. However, the confusion
matrix of the model, \textbf{Figure 3}, shows the errors are concentrated on the healthy control
categories, while categorization of diagnosed subjects was substantially better.
This implies major depressive subjects show statistically significant and
distinct patterns in their TMS responses in comparison to healthy controls, and
may be considered evidence for depression-induced differences in
neuroplasticity.

However, it is important to note that the proxy for neuroplasticity in the
double pulse paradigm is precisely the relative amplitude of double pulses with
respect to test pulses. The raw data lacks any measure of relative amplitude.
Besides, regardless of the fact that the model above suggests a critical
difference in the response patterns between subject groups, its accuracy is
still poor.

@@

We then trained the model first with the aggregation of the $\rho$ feature, then
with the aggregation of the $\delta$ feature, and then with the aggregation of
both.

With the inclusion of the $\rho$ feature increased accuracy by a factor of
$\approx 2.1$ to $72.6\%$. The inclusion of the $\delta$ feature had a similar
impact, increasing accuracy to $73.4\%$.

@@

The inclusion of both the $\delta$ and $\rho$ features computed over the
logarithmically transformed EMG peak-to-peak improved the model's accuracy.
Concretely, accuracy was increased to $86\%$, with the following confusion
matrix.

\subsection{Random forest after $\ln$ transforms and weighted averages}

Since the neural response to transcranial stimulations follows an exponential
distribution (see \textbf{Cite section}), we experimented with the inclusion of
the features above applied to the logarithmically transformed peak-to-peak EMG.

As stated earlier, action potentials evoked by transcranial magnetic
stimulations follow a Gamma distribution that is very close to the exponential.
(see \textbf{Statistics} section). Experimentation has shown that the inclusion
of weighted variants of the features defined above greatly improves the
performance of a random forest model (see \textbf{Empirical results}).

The use of weighted instead of arithmetic averages may be useful in dealing with
the excessive influence of outliers or highly spread out points in the feature.
For example, the weight vectors may be computed using the MAD or the
inverse-variance of each $x$.

As stated in section \textbf{3}, we experimented with using weighted
versions of our pulse-specific MRA. They consisted in weighting test pulses
with some appropriate weight vector. For example, including $\rho$ and
$\delta$ using inverse-variance weights, we obtained an accuracy of
$90.2\%$, with the following confusion matrix.

\begin{figure*}
    \centering
    \includegraphics[scale=0.4]{final-cm}
    \caption{Confusion matrix: $\rho(x)$, $\delta(x), \rho(\ln x), \delta(\ln
    x)$, $\rho_w(x)$ and $\delta_w$(x).}
    \label{fig:figure5}
\end{figure*}

\section{Discussion}

The results in \textbf{Section 6} demonstrate that pulse-specific MRA greatly
improve the accuracy of a random forest model trained over TMS data for subject
classification. More importantly, it becomes clear that, by using the engineered
features $\delta$ and $\rho$, researchers can attain evidence in favour or
against hypotheses pertaining to differences among subject groups. Particularly,
the model provides otherwise unattainable evidence in favour of sleep-modulated,
depression-induced differences in neuroplasticity.

@@

This paper shows that pulse-specific relative amplitude measures allow for the
use of making machine learning models over TMS experimental data. Such models
can play an important part in future research by introducing new ways of
both analyzing and extracting meaningful insights from TMS results.

@@

>>>>

\title{\textbf{Modelling Neuroplasticity: Using Machine Learning to Learn About TMS Data}}
\author{Santiago López Pereyra, Jennifer Goldschmied, Coauthor 3,
Coauthor 4, \ldots}
\maketitle

@@

The distribution of the evoked potentials was always exponential. However, the
$\beta$ parameter of the distribution varies across subject groups and session
types, as shown in Figures \ref{fig:figure1} and \ref{fig:figure2} and in the
table below:

@@

    \begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|  }
    \hline
    &       Mean & Median \\
    \hline
    HC BL & 196.77 & 149.36 \\
    \hline
    HC SWD & 165.98 & 99.76 \\
    \hline 
    MDD BL & 187.56 & 127.69\\ 
    \hline 
    MDD SWD & 247.57 & 151.07 \\
    \hline
    \end{tabular}

@@

Observe that each of the means in the above table corresponds to the estimated
$\beta$ parameter of the distribution of the paired pulses on each subject
group.

@@

In order to test these measures, we look at whether they improve the performance
of a ML classification model and in what degree.\footnote{We use a
classification model for practical reasons ---namely, because the samples in our
data were conveniently labeled by subject group and sleep session.} Our case
study is a random forest model in Python tasked with classifying every
observation with the appropriate subject label. The model we use is an AdaBoost
classifier consisting of 20 decision trees. Every tree in the ensemble has a
maximum depth of $8$. The boosting algorithm is \textit{SAMME.R}. The learning
rate is set to $\eta = 0.01$. We use holdout cross-validation to evaluate the
performance of the model. The total number of values in our dataset is $n =
9382$.

The goal of the classifier is to infer, based on the properties of each TMS
response, the group of the subject upon which the stimulation was elicited, as
well as the type of sleep session after which it occurred. Our hypothesis is
that pulse-specific measures of relative amplitude will significantly improve
the model's performance.

@@

The first random forest model is our control group. We trained it on the raw
data ---that is, without the inclusion of pulse-specific MRA. The model's
accuracy on the testing set was of $\approx34.2 \%$. The confusion matrix of the
model, \textbf{Figure 3.}\textit{a}, shows that errors concentrated towards
healthy control categories. Additionally, the categorization of diagnosed
subjects was poor, although substantially better than that of healthy controls.

Although the model almost never confuses an MDD subject with a healthy control,
this does not necessarily imply that the model is capturing key differences
between these groups. On the contrary, this is more likely a consequence of the
model's bias in favour of the MDD categories. This bias is probably caused by
the fact that the dataset has more MDD subjects than healthy controls. In
consequence, the model learns the rudimentary strategy of always guessing for
one of the MDD categories.

@@

The following models incorporated the $\rho$ feature, then the $\delta$ feature,
and then both, respectively. The inclusion of the $\rho$ feature increased
accuracy by a factor of $\approx 2.1$ to $72.6\%$. Subsequently, the inclusion
of the $\delta$ feature had a similar impact, increasing accuracy to $73.4\%$.
Lastly, the inclusion of both the $\delta$ and $\rho$ features improved accuracy
to $\approx 87.1\%$. The confusion matrix of the model trained with both $\rho$
and $\delta$ can be observed in \textbf{Figure 3.}\textit{b}.

@@

\subsection{Random forest with additional features}

Since the neural response to transcranial stimulations follows an exponential
distribution (see Figures 1 and 2), we tested the $\rho$ and $\delta$ functions
over the logarithmically transformed peak-to-peak EMG. This had a negligible
positive impact in the overall performance.

On the contrary, the inclusion of weighted variants of the features defined
above greatly improves the performance of the model. The use of weighted instead
of arithmetic averages in the formulas for $\delta$ and $\rho$ is useful in
dealing with the excessive influence of outliers or highly spread out points in
the feature. For example, the weight vectors may be computed using the MAD or
the inverse-variance of each $x$.

For example, using inverse-variance weights accuracy increased to $\approx
90.2\%$. The confusion matrix of this last model can be appreciated in
\textbf{Figure 4}.

\begin{figure}[H]
    \centering
    \subfigure[$\rho, \delta, \rho(\ln x), \delta(\ln x), \rho_w$ and $\delta_w$
    included.]{\includegraphics[width=0.48\textwidth]{final-cm}}
    \caption{Confusion matrix}
    \label{fig:figure4}
\end{figure}

In summary, our results demonstrate that pulse-specific MRA greatly improve the
accuracy of a random forest model trained over TMS data for subject
classification. More importantly, it becomes clear that, by using the engineered
features $\delta$ and $\rho$, researchers can attain evidence in favour or
against hypotheses pertaining to differences among subject groups. Particularly,
the model provides otherwise unattainable evidence in favour of sleep-modulated,
depression-induced differences in neuroplasticity.

@@

This paper shows that pulse-specific relative amplitude measures allow for the
use of machine learning models over TMS experimental data. Such models
can play an important part in TMS research by introducing new ways of
both analyzing and extracting meaningful insights from the experimental results.

@@

\end{multicols}

\section{References}

\begin{enumerate}
    \item Lefaucheur, JP. “Transcranial Magnetic Stimulation.” Handbook of clinical neurology. U.S. National Library of Medicine. Accessed April 14, 2023. https://pubmed.ncbi.nlm.nih.gov/31277876. 
    \item Orth, M, AH Snijders, and JC Rothwell. “The Variability of Intracortical Inhibition and Facilitation.” Clinical neurophysiology : official journal of the International Federation of Clinical Neurophysiology. U.S. National Library of Medicine. Accessed April 14, 2023. https://pubmed.ncbi.nlm.nih.gov/14652096/. 
    \item Rossini, PM, D Burke, LG Cohen, Z Daskalakis, R Di Iorio, V Di Lazzaro, F Ferreri, et al. “Non-Invasive Electrical and Magnetic Stimulation of the Brain, Spinal Cord, Roots and Peripheral Nerves: Basic Principles and Procedures for Routine Clinical and Research Application. an Updated Report from an I.F.C.N. Committee.” Clinical neurophysiology : official journal of the International Federation of Clinical Neurophysiology. U.S. National Library of Medicine. Accessed April 14, 2023. https://pubmed.ncbi.nlm.nih.gov/25797650/. 
        \item Wassermann, EM. “Variation in the Response to Transcranial
            Magnetic Brain Stimulation in the General Population.” Clinical
            neurophysiology : official journal of the International Federation
            of Clinical Neurophysiology. U.S. National Library of Medicine.
            Accessed April 14, 2023. \newline https://pubmed.ncbi.nlm.nih.gov/12088713/. 
\end{enumerate}

\pagebreak
